Structuring College Access: The Market Segment Model and College Board Geomarkets

### 2025-07-09 text/manuscript-full-KS.qmd
revisions rq2 LA

@fig-la-rq2-race presents the racial/ethnic composition of purchased student profiles. For the more than 15,000 prospects in the resulting student lists at the low PSAT score thresholds (1070-1180), South and South Central Los Angeles (n=1,031) and East Los Angeles (n=871) Geomarket pools stand out as having an overwhelming majority of students of color, 98% and 95% respectively. Relative to other Geomarkets in the Los Angeles metropolitan area, such patterns are attributed to both South and South Central Los Angeles (89%) and East Los Angeles (64%) having substantially greater proportions of Hispanic prosepcts, as well as a larger proportion of Black prospects for South and South Central Los Angeles (6%) and larger share of Asian prospects for East Los Angeles (30%). Even among higher PSAT score ranges, South and South Central Los Angeles remains the most prominent Geomarket with significant proportions of Hispanic and Black student prospects, while East Los Angeles stands out for its substantial share of Asian prospects.

@@

@fig-la-rq2-midSAT-race shows how much each Geomarket accounts for the share of prospects in the middle PSAT score range by race/ethnicity. South and South Central Los Angeles accounts for only 2% of all purchased prospect profiles but accounts for 7% of Black prospects. The Reiverside, San Bernardino, & Ontario Geomarket accounts for 17% of all prospects but 25% of Black prospects. For Hispanic prospects, South and South Central Los Angeles (8% versus 2% for all prospects) and Riverside, San Bernadino, and Ontario (24% versus 17% for all prospects) also contributed a disproportinately larger proportions of prospects relative to their contributions to the overall pool in the Los Angeles metropolitan area. For Asian students, the Glendale and Pasadena (17% of Asian prospects versus 14% for all prospects), East Los Angeles (11% of Asian prospects versus 5% of all prospects) and Covina and West Covina (17% of Asain prospects versus 11% of all prospects) Geomarkets contributed a disproportionate larger share of Asian prospects relative to their contributions to the overall prospect pool. For the more than 2,500 White prospects in this pool, the San Fernando Valley -- West (28% of White prospects versus 20% of all prospects) and the West Los Angeles and West Beach (12% of White prospects versus 7% for all prospects) Geomarkets contribute disproportionately larger shares.

These results demonstrate that following the @RN4982 strategy of filtering on high-income Geomarkets and excluding low-income Geomarkets -- such as the South and South Central Los Angeles Geomarket and the Riverside, San Bernadino, and Ontario Geomarket -- would result in the disproportionate exclusiong of Black and Hispanic students from the resulting prospect pools.Similarly, using a Geomarket filter excluding East Los Angeles would result in the loss of disproportionate shares of Asian students.

\[KARINA -- CHECK YOUR NUMBERS FOR THIS PARAGRAPH; KS: I DID; THEY ARE CORRECT FOR ORDERS 44820 AND 546946; AM I MISSING SOMETHING/WHY DO YOU THINK THEY ARE WRONG?\] @fig-la-rq2-midSAT-ses presents Geomarket contributions by first-generation college student status, demonstrating the disproportionately larger share of first-generation college students contributed to the overall pool by East Los Angeles, South and South Central Los Angeles, and Riverside, San Bernadino, and Ontario Geomarkets. For students whose parents did not attend college, East Los Angeles (16% versus 5% of all prospects), South and South Central Los Angeles (10% versus 2% of all prospects), and Riverside, San Bernardino, and Ontario (24% versus 17% of all prospects) Geomarkets provided a disproportionately larger contribution of first-generation college students. <!--#  KARINA -- I CUT THIS TEXT. OK? These three Geomarkets also made disproportionately larger contributions of first-generation college student prospects whose parents attended but did not complete their degree, as well as Hollywood & Wilshire (7% versus 5% of all prospects), Long Beach (7% versus 5% of all prospects), and Covina and West Covina (13% versus 10% of all prospects).  -->

@@

Lastly, @fig-la-rq2-midSAT-combo presents Geomarket contributions to prospect pools by both first-generation college student status and race/ethnicity in the Los Angeles metropolitan area. Across all Hispanic prospects in the middle range PSAT orders, 25% were first-generation college students whose parents did not attend college, 25% were first-generation college students whose parents attended some college but did not complete their degrees, and 50% were not first-generation college students. By contrast, for the 156 Hispanic prospects from the South and South Central Geomarket, 54% had parents who never attended college, 28% had parents who attended some college, and only 18% had parents with a BA. Similarly, Hispanic prospects from East Los Angeles, from Hollywood & Wilshire, and from Riverside, San Bernardino, and Ontario were much more likely to have parents without a BA compared to all Hispanic prospects in the Los Angeles metro area.

@@

Among all Asian prospects whose contact information was purchased, approximately 11% were first-generation college students whose parents did not attend college, 21% were first-generation college students whose parents attended some college but did not complete their degree, and 68% were not first-generation college students. @fig-la-rq2-midSAT-combo highlights the East Los Angeles Geomarket as a key contributor of Asian prospects (n = 251), a significant proportion of whom are first-generation college students (35%) whose parents did not attend college. The Hollywood & Wilshire Geomarket also contributed a modest number of Asian first-generation college students (some college) at a disproportionately larger rate (35%).

While the pool of Black prospects is relatively small (n=186) in @fig-la-rq2-midSAT-combo, some patterns emerge by first-generation status and Geomarket. Overall, approximately 5% of Black prospects were first-generation college students-no college, 20% were first-generation college students-some college, and 74% not first-generation). The South and South Central Los Angeles Geomarket stands out for contributing a disproportionately high share (21%) of Black first-generation college students whose parents did not attend college, despite representing a relatively smaller overall number. Additionally, the Long Beach Geomarket is notable for its higher proportion (40%) of Black first-generation students whose parents attended college but did not complete a degree.

@@

```{r fig-la-rq2-race, echo = FALSE, fig.align = 'center', fig.cap = "Racial/Ethnic Composition of Purchased Student Profiles by Geomarket, Los Angeles Area", out.width = "100%"}
knitr::include_graphics('./../assets/images/rq2_la_race.png')
#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
#![The enrollment funnel](assets/images/enrollment_funnel.png)
```

```{r fig-la-rq2-midSAT-race, echo = FALSE, fig.align = 'center', fig.cap = "Los Angeles Geomarket Contribution to Purchased Student profiles by Racial/ethnic group, Middle-Range SAT orders", out.width = "100%"}
knitr::include_graphics('./../assets/images/rq2_la_midSAT_race.png')
#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
#![The enrollment funnel](assets/images/enrollment_funnel.png)

@@

```{r fig-la-rq2-midSAT-ses, echo = FALSE, fig.align = 'center', fig.cap = "Los Angeles Geomarket Contribution to Purchased Student profiles by First-Generation Status, Middle-Range SAT orders", out.width = "100%"}
  knitr::include_graphics('./../assets/images/rq2_la_midSAT_ses.png')
#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
#![The enrollment funnel](assets/images/enrollment_funnel.png)
```

```{r fig-la-rq2-midSAT-combo, echo = FALSE, fig.align = 'center', fig.cap = "Los Angeles Geomarket Contribution to Purchased Student profiles by First-Generation Status, Middle-Range SAT orders", out.width = "100%"}
knitr::include_graphics('./../assets/images/rq2_la_midSAT_combo.png')
#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
#![The enrollment funnel](assets/images/enrollment_funnel.png)
```

>>>>

@fig-la-rq2-race presents the racial/ethnic composition of purchased student profiles. For the nearly 15,000 prospects in the resulting student lists at the low PSAT score thresholds (1070-1180), South and South Central Los Angeles (n=1,035) and East Los Angeles (n=1,076) Geomarket pools stand out as having an overwhelming majority of students of color, 98% and 95% respectively. Relative to other Geomarkets in the Los Angeles metropolitan area, such patterns are attributed to both South and South Central Los Angeles (88%) and East Los Angeles (61%) having substantially greater proportions of Hispanic prosepcts, as well as a larger proportion of Black prospects for South and South Central Los Angeles (5%) and larger share of Asian prospects for East Los Angeles (31%). Even among higher PSAT score ranges, South and South Central Los Angeles remains the most prominent Geomarket with significant proportions of Hispanic, Asian, and Black student prospects. <!--#  while East Los Angeles stands out for its substantial share of Asian prospects.-->

@@

@fig-la-rq2-midSAT-race shows how much each Geomarket accounts for the share of prospects in the middle PSAT score range by race/ethnicity. South and South Central Los Angeles accounts for only 3% of all purchased prospect profiles but accounts for 8% of Black prospects. The Riverside, San Bernardino, & Ontario Geomarket accounts for 24% of all prospects but 29% of Black prospects. For Hispanic prospects, South and South Central Los Angeles (9% versus 3% for all prospects) and Riverside, San Bernadino, and Ontario (30% versus 24% for all prospects) also contributed a disproportionately larger proportions of prospects relative to their contributions to the overall pool in the Los Angeles metropolitan area. For Asian students, the Glendale and Pasadena (20% of Asian prospects versus 15% for all prospects), East Los Angeles (14% of Asian prospects versus 7% of all prospects) and Covina and West Covina (12% of Asain prospects versus 9% of all prospects) Geomarkets contributed a disproportionate larger share of prospects relative to their contributions to the overall prospect pool. For the more than 2,300 White prospects in this pool, the San Fernando Valley - West (17% of White prospects versus 11% of all prospects), West Los Angeles and West Beach (9% of White prospects versus 6% for all prospects), and Glendale and Pasadena (17% of White prospects versus 15% for all prospects) Geomarkets contribute disproportionately larger shares.

These results demonstrate that following the @RN4982 strategy of filtering on high-income Geomarkets and excluding low-income Geomarkets -- such as the South and South Central Los Angeles Geomarket and the Riverside, San Bernardino, and Ontario Geomarket -- would result in the disproportionate exclusion of Black and Hispanic students from the resulting prospect pools. Similarly, using a Geomarket filter excluding East Los Angeles would result in the loss of disproportionate shares of Asian students.

@fig-la-rq2-midSAT-ses presents Geomarket contributions by first-generation college student status, demonstrating the disproportionately larger share of first-generation college students contributed to the overall pool by East Los Angeles, South and South Central Los Angeles, and Riverside, San Bernadino, and Ontario Geomarkets. For students whose parents did not attend college, East Los Angeles (18% versus 7% of all prospects), South and South Central Los Angeles (11% versus 3% of all prospects), and Riverside, San Bernardino, and Ontario (27% versus 24% of all prospects) Geomarkets provided a disproportionately larger contribution of first-generation college students. <!--#  KARINA -- I CUT THIS TEXT. OK? These three Geomarkets also made disproportionately larger contributions of first-generation college student prospects whose parents attended but did not complete their degree, as well as Hollywood & Wilshire (7% versus 5% of all prospects), Long Beach (7% versus 5% of all prospects), and Covina and West Covina (13% versus 10% of all prospects).  -->

@@

Lastly, @fig-la-rq2-midSAT-combo presents Geomarket contributions to prospect pools by both first-generation college student status and race/ethnicity in the Los Angeles metropolitan area. Across all Hispanic prospects in the middle range PSAT orders, 26% were first-generation college students whose parents did not attend college, 26% were first-generation college students whose parents attended some college but did not complete their degrees, and 49% were not first-generation college students. By contrast, for the 175 Hispanic prospects from the South and South Central Geomarket, 51% had parents who never attended college, 28% had parents who attended some college, and only 21% had parents with a BA. Similarly, Hispanic prospects from East Los Angeles and from Riverside, San Bernardino, and Ontario were more likely to have parents without a BA compared to all Hispanic prospects in the Los Angeles metro area.

@@

Among all Asian prospects whose contact information was purchased, approximately 11% were first-generation college students whose parents did not attend college, 22% were first-generation college students whose parents attended some college but did not complete their degree, and 67% were not first-generation college students. @fig-la-rq2-midSAT-combo highlights the East Los Angeles Geomarket as a key contributor of Asian prospects (n = 276), a significant proportion of whom are first-generation college students (34%) whose parents did not attend college. The Hollywood & Wilshire Geomarket also contributed a modest number of Asian first-generation college students (some college) at a disproportionately larger rate (34%).

While the pool of Black prospects is relatively small (n=182) in @fig-la-rq2-midSAT-combo, some patterns emerge by first-generation status and Geomarket. Overall, approximately 6% of Black prospects were first-generation college students-no college, 21% were first-generation college students-some college, and 74% not first-generation). The South and South Central Los Angeles Geomarket stands out for contributing a disproportionately high share (13%) of Black first-generation college students whose parents did not attend college, despite representing a relatively smaller overall number. Additionally, the Long Beach Geomarket is notable for its higher proportion (42%) of Black first-generation students whose parents attended college but did not complete a degree.

@@

```{r fig-la-rq2-race, echo = FALSE, fig.align = 'center', fig.cap = "Racial/Ethnic Composition of Purchased Student Profiles by Geomarket, Los Angeles Area", out.width = "100%"}
knitr::include_graphics('./../results/graphs/rq2/rq2_los_angeles_race_row_plot.png')

#knitr::include_graphics('./../assets/images/rq2_la_race.png') [RESULTS BASED ON ZIP GEOMARKETS]
#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
#![The enrollment funnel](assets/images/enrollment_funnel.png)
```

```{r fig-la-rq2-midSAT-race, echo = FALSE, fig.align = 'center', fig.cap = "Los Angeles Geomarket Contribution to Purchased Student profiles by Racial/ethnic group, Middle-Range SAT orders", out.width = "100%"}
knitr::include_graphics('./../results/graphs/rq2/rq2_los_angeles_race_col_plot_448420_546946.png')
#knitr::include_graphics('./../assets/images/rq2_la_midSAT_race.png')[RESULTS BASED ON ZIP GEOMARKETS]
#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
#![The enrollment funnel](assets/images/enrollment_funnel.png)

@@

```{r fig-la-rq2-midSAT-ses, echo = FALSE, fig.align = 'center', fig.cap = "Los Angeles Geomarket Contribution to Purchased Student profiles by First-Generation Status, Middle-Range SAT orders", out.width = "100%"}
knitr::include_graphics('./../results/graphs/rq2/rq2_los_angeles_firstgen_col_plot_448420_546946.png')
 # knitr::include_graphics('./../assets/images/rq2_la_midSAT_ses.png') [RESULTS BASED ON ZIP GEOMARKETS]
#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
#![The enrollment funnel](assets/images/enrollment_funnel.png)
```

```{r fig-la-rq2-midSAT-combo, echo = FALSE, fig.align = 'center', fig.cap = "Los Angeles Geomarket Contribution to Purchased Student profiles by First-Generation Status, Middle-Range SAT orders", out.width = "100%"}
knitr::include_graphics('./../results/graphs/rq2b/rq2b_los_angeles_order_448420_546946_race_by_firstgen.png')
#knitr::include_graphics('./../assets/images/rq2_la_midSAT_combo.png')[RESULTS BASED ON ZIP GEOMARKETS]
#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
#![The enrollment funnel](assets/images/enrollment_funnel.png)
```
